Easy as Pi, Volume II: On the Goodput and Energy of Network Coding Schemes in the Raspberry Pi

### 2016-08-15 measurements.tex
corrections and density regions estimate script
corrections

In Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, \ref{fig:enc_good_rasp1_gen_gf256},
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256};
it can be seen that the goodput trends of five codes as a function of
the generation size: \ac{RLNC}, \ac{SRLNC} with $d = [0.02,\ 0.1]$
and \ac{TSNC} with an extra parameter that we mention. We define
$C$\fxnote{C is already defined as the set of l.i. coded packets}
as the number of density regions in the \ac{TSNC} transmission
process. The maximum number of possible regions depends
on the generation size. See equation~\eqref{eq:feedback_dofs}.
The larger the generation, the more density regions may
be formed and more density changes are possible.
Throughout this paper, \ac{TSNC} is configured to use the maximum number
of density regions possible. 
As there can be at least $C=1$ density region in a transmission, we use
$C=0$ to indicate the maximum possible density regions. This is
used for \ac{TSNC} in plots where the generation size is not fixed.
%In fact, \ac{TSNC} is configured to use the
%maximum possible density regions in all plots throughout this paper.
%The larger the generation, the more regions are formed and
%more density changes are possible.
%Throughout the section, we evaluate \ac{TSNC} with as many possible
%density regions that are possible to form during a transmission. This number
%grows as the generation increases, so we define $C=0$ as the maximum
%number of possible regions.
%
%We first evaluate \ac{TSNC} with no
%changes, e.g. $C = 0$ and the two amounts for dependent packets mentioned
%earlier. This means that we tolerate at most 8 or 16\fxnote{dependent packets} packets regardless of
%the generation. The goal with setting first \ac{TSNC} with these values
%is to observe first the sole effect of the budget in the code since
%$C = 0$ implies no density changes during the whole process.
%
For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the lowest performance in terms of goodput and
\ac{TSNC} with $e = 16$, the highest regardless of the \ac{Raspi} model.
Given that the processing time depends in the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets\fxnote{The encoding process speed depends on the total number of original packets to mix. Thus coding density and number of coded packets per generation}. Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of \ac{l.d.} packets is. So, basically
the sparser the codes, then more overhead due to transmissions of \ac{l.d.}
packets since it is more frequent to generate similar set of packets.
Excluding the coding coefficients overhead, the overhead due to
transmissions of \ac{l.d.} packets might be high for the sparse schemes.
For example, if we consider \ac{TSNC} with $e = 16$ and $g = 16$
in Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, the budget in this case permits
to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the overhead from appending
the coding coefficients. This happens because \ac{TSNC}
has been allowed to add too much redundancy in this case. Given that $C=0$,
this permits it to use a single very low code density $d$. For \ac{RLNC},
this is not the case since the occurrence of \ac{l.d.} coded packets is
small, because all coding coefficients are used. Even for $GF(2)$, the
average amount of redundant packets for \ac{RLNC} has proven to be 1.6
packets after $g$ have been transmitted \cite{trullols2011exact,zhao2012notes},
but less than the cases where sparse codes are utilized. Overall, we
observe that there is a trade-off between goodput and \ac{l.d.} coded
packets transmission overhead.

@@

In Fig.~\ref{fig:dec_good_rasp1} and \ref{fig:dec_good_rasp2}, we
observe the same generation and packet size effects reported in the
encoding case. However, we do observe in
Fig.~\ref{fig:dec_good_rasp1_gen_gf2}, \ref{fig:dec_good_rasp1_gen_gf256},
\ref{fig:dec_good_rasp2_gen_gf2} and \ref{fig:dec_good_rasp2_gen_gf256};
that duplicating\fxnote{What does duplicating the generation size mean? Did you mean doubling?} the generation size does not reduce the goodput by a
factor of four. In principle, given that Gaussian elimination scales as
$\mathcal{O}(g^3)$ (and thus the processing time), we would expect that
the goodput to scale as $\mathcal{O}(R_{proc,dec}) = \mathcal{O}(g/g^{3}) =
\mathcal{O}(g^{-2})$. This would imply that duplicating\fxnote{doubling?} the generation
size, should reduce the goodput by a factor of four, which is not the case. Instead,
the goodput is only reduced by a factor of two. This is only possible if
the Gaussian elimination is $\mathcal{O}(g^2)$. A study in
\cite{paramanathan2013leanandmean} for \ac{RLNC} speeds in commercial
devices indicated that this is effectively the case. The reason is that
the $g^2$ scaling factor in the scaling law\fxnote{scaling estimate or scaling expression or something else instead of scaling law} of the Gaussian elimination
is much
higher than the $g^3$ scaling for $g < 512$. Particularly, this factor
relates to the number of field elements in a packet size as mentioned
in \cite{paramanathan2013leanandmean}.

>>>>

In Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, \ref{fig:enc_good_rasp1_gen_gf256},
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256};
it can be seen that the goodput trends of five codes as a function of
the generation size: \ac{RLNC}, \ac{SRLNC} with $d = [0.02,\ 0.1]$
and \ac{TSNC} with an extra parameter that we mention. We define
$C$\fxnote{C is already defined as the set of l.i. coded packets}
as the number of density regions in the \ac{TSNC} transmission
process. The maximum number of possible regions depends
on the generation size. See~\eqref{eq:feedback_dofs}.
The larger the generation, the more density regions may
be formed and more density changes are possible.
Throughout this paper, \ac{TSNC} is configured to use the maximum number
of density regions possible. 
As there can be at least $C=1$ density region in a transmission, we use
$C=0$ to indicate the maximum possible density regions. This is
used for \ac{TSNC} in plots where the generation size is not fixed.
%In fact, \ac{TSNC} is configured to use the
%maximum possible density regions in all plots throughout this paper.
%The larger the generation, the more regions are formed and
%more density changes are possible.
%Throughout the section, we evaluate \ac{TSNC} with as many possible
%density regions that are possible to form during a transmission. This number
%grows as the generation increases, so we define $C=0$ as the maximum
%number of possible regions.
%
%We first evaluate \ac{TSNC} with no
%changes, e.g. $C = 0$ and the two amounts for dependent packets mentioned
%earlier. This means that we tolerate at most 8 or 16\fxnote{dependent packets} packets regardless of
%the generation. The goal with setting first \ac{TSNC} with these values
%is to observe first the sole effect of the budget in the code since
%$C = 0$ implies no density changes during the whole process.
%
For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the lowest performance in terms of goodput and
\ac{TSNC} with $e = 16$, the highest regardless of the \ac{Raspi} model.
Given that the processing time depends in the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets\fxnote{The encoding process speed depends on the total number of original packets to mix. Thus coding density and number of coded packets per generation}. Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of \ac{l.d.} packets is. So, basically
the sparser the codes, then more overhead due to transmissions of \ac{l.d.}
packets since it is more frequent to generate similar set of packets.
Excluding the coding coefficients overhead, the overhead due to
transmissions of \ac{l.d.} packets might be high for the sparse schemes.
For example, if we consider \ac{TSNC} with $e = 16$ and $g = 16$
in Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, the budget in this case permits
to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the overhead from appending
the coding coefficients. This happens because \ac{TSNC}
has been allowed to add too much redundancy in this case. Given that $C=0$,
this permits it to use a single very low code density $d$. For \ac{RLNC},
this is not the case since the occurrence of \ac{l.d.} coded packets is
small, because all coding coefficients are used. Even for $GF(2)$, the
average amount of redundant packets for \ac{RLNC} has proven to be 1.6
packets after $g$ have been transmitted \cite{trullols2011exact,zhao2012notes},
but less than the cases where sparse codes are utilized. Overall, we
observe that there is a trade-off between goodput and \ac{l.d.} coded
packets transmission overhead.

@@

In Fig.~\ref{fig:dec_good_rasp1} and \ref{fig:dec_good_rasp2}, we
observe the same generation and packet size effects reported in the
encoding case. However, we do observe in
Fig.~\ref{fig:dec_good_rasp1_gen_gf2}, \ref{fig:dec_good_rasp1_gen_gf256},
\ref{fig:dec_good_rasp2_gen_gf2} and \ref{fig:dec_good_rasp2_gen_gf256};
that doubling the generation size does not reduce the goodput by a
factor of four. In principle, given that Gaussian elimination scales as
$\mathcal{O}(g^3)$ (and thus the processing time), we would expect that
the goodput to scale as $\mathcal{O}(R_{proc,dec}) = \mathcal{O}(g/g^{3}) =
\mathcal{O}(g^{-2})$. This would imply that doubling the generation
size, should reduce the goodput by a factor of four, which is not the case. Instead,
the goodput is only reduced by a factor of two. This is only possible if
the Gaussian elimination is $\mathcal{O}(g^2)$. A study in
\cite{paramanathan2013leanandmean} for \ac{RLNC} speeds in commercial
devices indicated that this is effectively the case. The reason is that
the $g^2$ scaling factor in the scaling law\fxnote{scaling estimate or scaling expression or something else instead of scaling law} of the Gaussian elimination
is much
higher than the $g^3$ scaling for $g < 512$. Particularly, this factor
relates to the number of field elements in a packet size as mentioned
in \cite{paramanathan2013leanandmean}.

### 2016-08-15 metrics.tex
corrections

\subsubsection{Average Power Expenditure}
To extract the average power for a given configuration in our setup, we
first calculate the average current in each\fxnote{each of?} the idle state $I_{idle,avg}$
and the processing state $I_{proc,avg}$ for all the available sample for
the given configuration. Regardless of the current type, the average value
of the sample set of a given number of samples $N_s$ is:
%
\begin{align} \label{eq:avg_current}
I_{avg} = \frac{1}{N_s}\sum_{k=1}^{N_s} I_{k}  ~[\mathrm{Ampere}]
\end{align}
%
With the average current from \eqref{eq:avg_current}, we compute the average
current used for processing with respect to the idle state, by substracting
$I_{idle,avg}$ from $I_{proc,avg}$. Then, the result is multiplied by the
supply voltage to obtain the average power during the considered configuration,
given as:
%
\begin{align} \label{eq:avg_power}
P_{avg} = V_{supply}(I_{proc,avg} - I_{idle,avg}) ~[\mathrm{Watt}]
\end{align}
%
\subsubsection{Energy per Bit Consumption}
To get this metric for a given configuration, we express the energy
as the product of the average power by the processing time $T_{proc}
~[\mathrm{sec/byte}]$ obtained from the goodput measurement for the same
configuration. In this way, we can relate the processing time with the
goodput, the packet and the generation size as shown:
%
\begin{align} \label{eq:energy_per_bit}
E_b = P_{avg} T_{proc,bit} = P_{avg} \times \frac{T_{proc}}{8gB} = \frac{P_{avg}}{8R_{proc}} ~[\mathrm{Joule}]
\end{align}
%

>>>>

\subsubsection{Average Power Expenditure}
To extract the average power for a given configuration in our setup, we
first calculate the average current in the idle state $I_{idle,avg}$
and the processing state $I_{proc,avg}$ for all the available sample for
the given configuration. Regardless of the current type, the average value
of the sample set of a given number of samples $N_s$ is:
%
\begin{align} \label{eq:avg_current}
I_{avg} = \frac{1}{N_s}\sum_{k=1}^{N_s} I_{k}  ~[\mathrm{Ampere}]
\end{align}
%
With the average current from \eqref{eq:avg_current}, we compute the average
current used for processing with respect to the idle state, by substracting
$I_{idle,avg}$ from $I_{proc,avg}$. Then, the result is multiplied by the
supply voltage to obtain the average power during the considered configuration,
given as:
%
\begin{align} \label{eq:avg_power}
P_{avg} = V_{supply}(I_{proc,avg} - I_{idle,avg}) ~[\mathrm{Watt}]
\end{align}
%
\subsubsection{Energy per Bit Consumption}
To get this metric for a given configuration, we express the energy
as the product of the average power by the processing time $T_{proc}
~[\mathrm{sec/byte}]$ obtained from the goodput measurement for the same
configuration. In this way, we can relate the processing time with the
goodput, the packet and the generation size as shown:
%
\begin{align} \label{eq:energy_per_bit}
E_b = P_{avg} T_{proc,bit} = P_{avg} \times \frac{T_{proc}}{8gB} = \frac{P_{avg}}{8R_{proc}} ~[\mathrm{Joule}]
\end{align}
%

### 2016-08-15 schemes.tex
corrections

\ac{TSNC} can be considered an extension to \ac{SRLNC}. The key idea is
not only for the encoder to generate sparse coded packets, but also to
modify the code density of the packets progressively as required. As a
decoder accomulates more coded packets, the probability that the next
received coded packet will be \ac{l.d.} increases~\cite{feizi2014tunable}.
\textcolor{red}{We can cite Feizi2012 here. That is the paper with the markov chain}

>>>>

\ac{TSNC} can be considered an extension to \ac{SRLNC}. The key idea is
not only for the encoder to generate sparse coded packets, but also to
modify the code density of the packets progressively as required. As a
decoder accomulates more coded packets, the probability that the next
received coded packet will be \ac{l.d.} increases~\cite{Feizi2012}.
%\cite{feizi2014tunable}.
